---
title: "C183 Project 4"
author: "Huimin Zhang"
date: "4/21/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## (a) Use your project data in the period 01-Jan-2013 to 01-Jan-2018.
### (1) Compute estimates for alpha, beta, and mse, by regressing each stock’s return on the S&P 500

```{r}
#Read csv file:
a <- read.csv("stockData.csv", sep=",", header=TRUE)

#Convert adjusted close prices into returns:
r <- (a[-1,3:ncol(a)]-a[-nrow(a),3:ncol(a)])/a[-nrow(a),3:ncol(a)]

#Compute the betas:
covmat <- var(r)
beta <- covmat[1,-1] / covmat[1,1]

#Keep only the stocks with positive betas:
rrr <- r[, -c(1, which(beta<0)+1)]

#Initialize
beta <- rep(0, ncol(rrr))
alpha <- rep(0, ncol(rrr))
mse <- rep(0, ncol(rrr))
stock <- rep(0, ncol(rrr))
var_beta <- rep(0, ncol(rrr))

for(i in 1:ncol(rrr)){
  q <- lm(data=rrr, formula=rrr[,i] ~ r[,1])
  stock[i] <- i
  alpha[i] <- q$coefficients[1] 
	beta[i] <- q$coefficients[2] 
  mse[i] <- summary(q)$sigma^2
  var_beta[i] <- vcov(q)[2,2]
}

#Matrix of alpha, beta, and mse
xx <- (cbind(stock, alpha, beta, mse))
xx
```

### (2) Construct the 30 × 30 variance covariance matrix based on the single index model.

```{r}
covmat_sim <- covmat[1,1] * (beta %*% t(beta)) + diag(mse)
covmat_sim[1:5, 1:5]
```

### (3) Draw the frontier on the same plot as in project 2(e).

```{r, warning=FALSE}
## The frontiers using the historical variance covariance matrix (project 2)

#Convert adjusted close prices into returns:
rr <- (a[-1,4:ncol(a)] - a[-nrow(a),4:ncol(a)]) / a[-nrow(a),4:ncol(a)]

#Compute the means:
means.rr <- colMeans(rr)
#Find the covariance matrix:
covmat.rr <- cov(rr)
#Compute the vector of variances:
variances.rr <- diag(covmat.rr)

ones <- matrix(rep(1, 30))

#Compute A:
A <- as.numeric(t(ones) %*% solve(covmat.rr) %*% means.rr)
#Compute B:
B <- as.numeric(t(means.rr) %*% solve(covmat.rr) %*% means.rr)

#Compute C:
C <- as.numeric(t(ones) %*% solve(covmat.rr) %*% ones)

#Compute D:
D <- as.numeric(B*C - A^2)

plot(0, A/C, main = "Portfolio possibilities curve (hyperbola)", type = "n", 
     xlab = "Risk (standard deviation)", ylab = "Expected Return", 
     xlim = c(-2*sqrt(1/C), 4*sqrt(1/C)), ylim = c(-2*A/C, 4*A/C))

#Plot center of the hyperbola:
points(0, A/C, pch = 19)
text(0, A/C+0.005, "historical variance model")

#Plot transverse and conjugate axes:
abline(v = 0, lty = 2) #Also this is the y-axis.
abline(h = A/C)

#Plot the x-axis:
abline(h = 0, lty = 2)

#Find the asymptotic:
V <- seq(-1, 1, 0.001)
A1 <- A/C + V * sqrt(D/C)
A2 <- A/C - V * sqrt(D/C)
points(V, A1, type = "l")
points(V, A2, type = "l")

#Efficient frontier:
minvar <- 1/C
minE <- A/C
sdeff <- seq((minvar)^0.5, 1, by = 0.0001)
options(warn = -1)
y1 <- minE + sqrt(D*(C*sdeff^2 - 1)) * minvar
y2 <- minE - sqrt(D*(C*sdeff^2 - 1)) * minvar
options(warn = 0)

points(sdeff, y1, type = "l")
points(sdeff, y2, type = "l")

## The frontiers using the variance covariance matrix with inputs from the single index model.
#Compute A:
A_sim <- as.numeric(t(ones) %*% solve(covmat_sim) %*% means.rr)
#Compute B:
B_sim <- as.numeric(t(means.rr) %*% solve(covmat_sim) %*% means.rr)

#Compute C:
C_sim <- as.numeric(t(ones) %*% solve(covmat_sim) %*% ones)

#Compute D:
D_sim <- as.numeric(B_sim*C_sim - A_sim^2)

#Plot center of the hyperbola:
points(0, A_sim/C_sim, pch = 19, col = "blue")
text(0, A_sim/C_sim-0.005, "single index model", col = "blue")

#Find the asymptotic:
V <- seq(-1, 1, 0.001)
A1_sim <- A_sim/C_sim + V * sqrt(D_sim/C_sim)
A2_sim <- A_sim/C_sim - V * sqrt(D_sim/C_sim)
points(V, A1_sim, type = "l", col = "blue")
points(V, A2_sim, type = "l", col = "blue")

#Efficient frontier:
minvar_sim <- 1/C_sim
minE_sim <- A_sim/C_sim
sdeff_sim <- seq((minvar_sim)^0.5, 1, by = 0.000001)
options(warn = -1)
y1_sim <- minE_sim + sqrt(D_sim*(C_sim*sdeff_sim^2 - 1)) * minvar_sim
y2_sim <- minE_sim - sqrt(D_sim*(C_sim*sdeff_sim^2 - 1)) * minvar_sim
options(warn = 0)

points(sdeff_sim, y1_sim, type = "l", col = "blue")
points(sdeff_sim, y2_sim, type = "l", col = "blue")
```

## (b) Adjusting the betas using Blume’s and Vasicek’s techniques.
### For the Blume technique
```{r}
a1 <- read.csv("stockData_test.csv", sep=",", header=TRUE)

#Convert adjusted close prices into returns:
r1 <- (a1[-1,3:ncol(a1)]-a1[-nrow(a1),3:ncol(a1)])/a1[-nrow(a1),3:ncol(a1)]

#Compute the variance covariance matrix of the returns for each period:
covmat1 <- var(r1)

#Compute the betas in each period:
beta1 <- covmat1[1,-1] / covmat1[1,1]

#Here is the plot of the betas in period 2 against the betas in period 1:
plot(beta, beta1)

#Correlation between the betas in the two periods:
cor(beta, beta1)

#Adjust betas using the Blume's technique:
q1 <- lm(beta1 ~ beta)

beta1_adj_blume <- q1$coef[1] + q1$coef[2]*beta1
beta1_adj_blume
```

### For the Vasicek technique
```{r}
#Adjusting the betas using the Vasicek's technique:
beta_adj_vasicek <- var_beta*mean(beta)/(var(beta)+var_beta) + var(beta)*beta/(var(beta)+var_beta)
beta_adj_vasicek

PRESS <- sum((beta_adj_vasicek - beta1)^2) / 30
PRESS
```
