---
title: "C183 Project 1"
author: "Huimin Zhang"
date: "4/2/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## (a) (b) Import the data in R and convert the adjusted close prices into returns.

```{r}
#Read csv file:
a <- read.csv("stockData.csv", sep=",", header=TRUE)
dim(a)
a[1:5, 1:5]

#Convert adjusted close prices into returns:
r <- (a[-1,3:ncol(a)]-a[-nrow(a),3:ncol(a)])/a[-nrow(a),3:ncol(a)]
r[1:5, 1:5]
```

## (c) Compute the means of the 31 assets, the standard deviations, and the variance covariance matrix.

```{r}
#Compute mean vector:
means <- colMeans(r)
means[1:5]

#Compute variance covariance matrix:
covmat <- cov(r)
covmat[1:5, 1:5]

#Compute correlation matrix:
cormat <- cor(r)
cormat[1:5, 1:5]

#Compute the vector of variances:
variances <- diag(covmat)
variances[1:5]

#Compute the vector of standard deviations:
stdev <- diag(covmat)^.5
stdev[1:5]
```

## (d) Plot the 31 assets on the space expected return against standard deviation.

```{r}
plot(stdev, means, xlab = "Standard deviation", ylab = "Expected return", 
     main = "Plot of expected return against\n standard deviation of the 31 assets")
```

## (e) Compute the mean and standard deviation of this portfolio and add it on the plot of question (c).

```{r}
#Convert adjusted close prices into returns:
rr <- (a[-1,4:ncol(a)]-a[-nrow(a),4:ncol(a)])/a[-nrow(a),4:ncol(a)]

#Compute the means:
means.rr <- colMeans(rr)
means.rr[1:5]

#Find the covariance matrix:
covmat.rr <- cov(rr)
covmat.rr[1:5, 1:5]

#Equal weights:
X <- rep(1/ncol(rr), ncol(rr))

#Expected return:
ER <- sum(X * means.rr)
ER

#Risk(standard deviation):
SD <- sqrt(t(X) %*% covmat.rr %*% X)
SD

#Add it on the plot of question (c).
plot(stdev, means, xlab = "Standard deviation", ylab = "Expected return", 
     main = "Plot of expected return against\n standard deviation of the 31 assets",
     xlim = c(0,0.2))
points(SD, ER, col = "red", pch = 19)
text(SD, ER+0.0025, "equal allocation", col = "red")
```

## (f) Add on the plot the minimum risk portfolio.

```{r}
#Minimum Weight
one <- matrix(rep(1, 30))
X_min_n <- solve(covmat.rr) %*% one 
X_min_d <- as.vector(t(one) %*% solve(covmat.rr) %*% one)
X_min <- X_min_n / X_min_d
X_min[1:5]

#Expected return:
ER_min <- t(X_min) %*% means.rr
ER_min

#Risk(standard deviation):
SD_min <- sqrt(t(X_min) %*% covmat.rr %*% X_min)
SD_min

plot(stdev, means, xlab = "Standard deviation", ylab = "Expected return", 
     main = "Plot of expected return against\n standard deviation of the 31 assets",
     xlim = c(0,0.2))
points(SD, ER, col = "red", pch = 19)
text(SD, ER+0.0025, "equal allocation", col = "red")
points(SD_min, ER_min, col = "blue", pch = 19)
text(SD_min, ER_min+0.0025, "minimum risk", col = "blue")
```

