---
title: "C183 Project 2"
author: "Huimin Zhang"
date: "4/6/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## (a) Compute A, B, C, D.

```{r}
#Read csv file:
a <- read.csv("stockData.csv", sep=",", header=TRUE)
dim(a)
a[1:5, 1:5]

#Convert adjusted close prices into returns:
rr <- (a[-1,4:ncol(a)] - a[-nrow(a),4:ncol(a)]) / a[-nrow(a),4:ncol(a)]
rr[1:5, 1:5]

#Compute the means:
means.rr <- colMeans(rr)

#Find the covariance matrix:
covmat.rr <- cov(rr)

#Compute the vector of variances:
variances.rr <- diag(covmat.rr)

#Compute the vector of standard deviations:
stdev.rr <- diag(covmat.rr)^.5

ones <- matrix(rep(1, 30))

#Compute A:
A <- as.numeric(t(ones) %*% solve(covmat.rr) %*% means.rr)
A

#Compute B:
B <- as.numeric(t(means.rr) %*% solve(covmat.rr) %*% means.rr)
B

#Compute C:
C <- as.numeric(t(ones) %*% solve(covmat.rr) %*% ones)
C

#Compute D:
D <- as.numeric(B*C - A^2)
D
```

## (b) Compute the values of the two Lagrange multipliers.

```{r}
E <- mean(means.rr)
E

lambda1 <- (C*E - A) / D
lambda1

lambda2 <- (B - A*E) / D
lambda2
```

## (c) Find the composition of the efficient portfolio given the return E.

```{r}
E_g <- 0.05

lambda1_g <- (C*E_g - A) / D
lambda1_g

lambda2_g <- (B - A*E_g) / D
lambda2_g

X_g <- solve(covmat.rr) %*% (lambda1_g * means.rr + lambda2_g * ones)
X_g
```

## (d) Plot the frontier in the mean-variance space (parabola)

```{r}
#Give values for E:
E_pb <- seq(-2, 2, 0.01)

#Compute sigma2 as a function of A,B,C,D, and E:
sigma2_pb <- (C*E_pb^2 - 2*A*E_pb + B) / D

#Plot E against sigma2:
plot(sigma2_pb, E_pb, type = "l", main = "Portfolio possibilities curve (parabola)",
     xlab = "Risk (variance)", ylab = "Expected Return")
```

## (e) Plot the frontier in the mean-standard deviation space using the hyperbola method.

```{r}
plot(0, A/C, main = "Portfolio possibilities curve (hyperbola)", type = "n", 
     xlab = "Risk (standard deviation)", ylab = "Expected Return", 
     xlim = c(-2*sqrt(1/C), 4*sqrt(1/C)), ylim = c(-2*A/C, 4*A/C))

#Plot center of the hyperbola:
points(0, A/C, pch = 19)

#Plot transverse and conjugate axes:
abline(v = 0, lty = 2) #Also this is the y-axis.
abline(h = A/C)

#Plot the x-axis:
abline(h = 0, lty = 2)

#Find the asymptotic:
V <- seq(-1, 1, 0.001)
A1 <- A/C + V * sqrt(D/C)
A2 <- A/C - V * sqrt(D/C)
points(V, A1, type = "l")
points(V, A2, type = "l")

#Efficient frontier:
minvar <- 1/C
minE <- A/C
sdeff <- seq((minvar)^0.5, 1, by = 0.0001)
options(warn = -1)
y1 <- minE + sqrt(D*(C*sdeff^2 - 1)) * minvar
y2 <- minE - sqrt(D*(C*sdeff^2 - 1)) * minvar
options(warn = 0)

points(sdeff, y1, type = "l")
points(sdeff, y2, type = "l")
```

## (f) On the plot in (e) add the 30 stocks, the S&P500, the equal allocation portfolio, the minimum risk portfolio, and the portfolio in (c).

```{r}
plot(0, A/C, main = "Portfolio possibilities curve (hyperbola)", type = "n", 
     xlab = "Risk (standard deviation)", ylab = "Expected Return", 
     xlim = c(-2*sqrt(1/C), 5*sqrt(1/C)), ylim = c(-3*A/C, 5*A/C))

#Plot center of the hyperbola:
points(0, A/C, pch = 19)

#Plot transverse and conjugate axes:
abline(v = 0, lty = 2) #Also this is the y-axis.
abline(h = A/C)

#Plot the x-axis:
abline(h = 0, lty = 2)

#Find the asymptotic:
V <- seq(-1, 1, 0.001)
A1 <- A/C + V * sqrt(D/C)
A2 <- A/C - V * sqrt(D/C)
points(V, A1, type = "l")
points(V, A2, type = "l")

#Efficient frontier:
minvar <- 1/C
minE <- A/C
sdeff <- seq((minvar)^0.5, 1, by = 0.0001)
options(warn = -1)
y1 <- minE + sqrt(D*(C*sdeff^2 - 1)) * minvar
y2 <- minE - sqrt(D*(C*sdeff^2 - 1)) * minvar
options(warn = 0)

points(sdeff, y1, type = "l")
points(sdeff, y2, type = "l")

#add the 30 stocks
points(stdev.rr, means.rr)

#add the S&P500
rsp <- (a[-1,3]-a[-nrow(a),3])/a[-nrow(a),3]
meansp <- mean(rsp)
sdsp <- sd(rsp)

points(sdsp, meansp, col = "orange", pch = 19)
text(sdsp, meansp-0.005, "S&P500", col = "orange")

#add the equal allocation portfolio
#Equal weights:
X <- rep(1/ncol(rr), ncol(rr))

#Expected return:
ER <- t(X) %*% means.rr

#Risk(standard deviation):
SD <- sqrt(t(X) %*% covmat.rr %*% X)

points(SD, ER, col = "red", pch = 19)
text(SD, ER+0.005, "equal allocation", col = "red")

#add the minimum risk portfolio
#Minimum risk weight
one <- matrix(rep(1, 30))
X_min_n <- solve(covmat.rr) %*% one 
X_min_d <- as.vector(t(one) %*% solve(covmat.rr) %*% one)
X_min <- X_min_n / X_min_d

#Expected return:
ER_min <- t(X_min) %*% means.rr

#Risk(standard deviation):
SD_min <- sqrt(t(X_min) %*% covmat.rr %*% X_min)

points(SD_min, ER_min, col = "blue", pch = 19)
text(SD_min, ER_min+0.005, "minimum risk", col = "blue")

#add the portfolio in (c).
#Expected return:
ER_g <- t(X_g) %*% means.rr

#Risk(standard deviation):
SD_g <- sqrt(t(X_g) %*% covmat.rr %*% X_g)

points(SD_g, ER_g, col = "green", pch = 19)
text(SD_g, ER_g-0.005, "prescribed expected return", col = "green")
```


## (g) Add three arbitrary portfolios on the plot of (c).

```{r}
set.seed(123)

xa <- runif(30)
x1 <- xa / sum(xa)
head(x1)

#Expected return:
ER1 <- t(x1) %*% means.rr

#Risk(standard deviation):
SD1 <- sqrt(t(x1) %*% covmat.rr %*% x1)

xb <- runif(30)
x2 <- xb / sum(xb)
head(x2)

#Expected return:
ER2 <- t(x2) %*% means.rr

#Risk(standard deviation):
SD2 <- sqrt(t(x2) %*% covmat.rr %*% x2)

xc <- runif(30)
x3 <- xc / sum(xc)
head(x3)

#Expected return:
ER3 <- t(x3) %*% means.rr

#Risk(standard deviation):
SD3 <- sqrt(t(x3) %*% covmat.rr %*% x3)

#the plot of (c)
#Plot the minimum risk portfolio:
plot(sqrt(1/C), A/C, pch=19, col="blue",
     xlim = c(0, 5*sqrt(1/C)), ylim = c(-3*A/C, 5*A/C),
     main = "Portfolio possibilities curve",
     xlab = "Risk (standard deviation)", ylab = "Expected Return")
    
#Efficient frontier:
minvar <- 1/C
minE <- A/C
sdeff <- seq((minvar)^0.5, 1, by = 0.0001)
options(warn = -1)
y1 <- minE + sqrt(D*(C*sdeff^2 - 1)) * minvar
y2 <- minE - sqrt(D*(C*sdeff^2 - 1)) * minvar
options(warn = 0)

points(sdeff, y1, type = "l")
points(sdeff, y2, type = "l")

#Expected return:
ER_g <- t(X_g) %*% means.rr

#Risk(standard deviation):
SD_g <- sqrt(t(X_g) %*% covmat.rr %*% X_g)

points(SD_g, ER_g, col = "green", pch = 19)
text(SD_g, ER_g-0.005, "prescribed expected return", col = "green")

points(SD1, ER1, col = "red", pch = 19)
text(SD1, ER1+0.005, "1", col = "red")

points(SD2, ER2, col = "red", pch = 19)
text(SD2, ER2+0.005, "2", col = "red")

points(SD3, ER3, col = "red", pch = 19)
text(SD3, ER3+0.005, "3", col = "red")
```
