---
title: "C183 Project 3"
author: "Huimin Zhang"
date: "4/15/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## (a) Convert the prices into returns for all the 5 stocks.

```{r}
a <- read.table("http://www.stat.ucla.edu/~nchristo/statistics_c183_c283/statc183c283_5stocks.txt", 
                header=T)
head(a)

#Convert adjusted close prices into returns:
r <- (a[-nrow(a),2:ncol(a)]-a[-1,2:ncol(a)])/a[-1,2:ncol(a)]
r[1:5, 1:5]
```

## (b) Compute the mean return for each stock and the variance-covariance matrix.

```{r}
#Compute mean vector:
means <- colMeans(r)
means

#Compute variance covariance matrix:
covmat <- cov(r)
covmat
```

## (c) Use only Exxon-Mobil and Boeing stocks: For these 2 stocks find the composition, expected return, and standard deviation of the minimum risk portfolio.

```{r}
#composition
x1_min <- (covmat[5,5] - covmat[1,5]) / (covmat[1,1] + covmat[5,5] - 2 * covmat[1,5])
x1_min
x2_min <- 1 - x1_min
x2_min

#expected return
r_min <- x1_min * as.numeric(means[1]) + x2_min * as.numeric(means[5])
r_min

#standard deviation
sd_min <- sqrt(x1_min^2 * covmat[1,1] + x2_min^2 * covmat[5,5] + 2 * x1_min * x2_min * covmat[1,5])
sd_min
```

## (d) Plot the portfolio possibilities curve and identify the efficient frontier on it.

```{r}
#many composition
xa_min <- seq(0, 1, 0.01)
xb_min <- 1 - xa_min

rp_min <- xa_min * as.numeric(means[1]) + xb_min * as.numeric(means[5])
sdp_min <- sqrt(xa_min^2 * covmat[1,1] + xb_min^2 * covmat[5,5] + 2 * xa_min * xb_min * covmat[1,5])

plot(sdp_min, rp_min, type = "l", xlab = "Risk (standard deviation)", ylab = "Expected Return", 
main = "Portfolio possibilities curve of Exxon-Mobil and Boeing")

#efficient frontier:
frontier <- data.frame(sdp_min, rp_min)
points(frontier[frontier$rp_min > r_min, ], col = "blue", type="l")
```

## (e) Use only Exxon-Mobil, McDonalds and Boeing stocks and assume short sales are allowed to answer the following question: For these 3 stocks compute the expected return and standard deviation for many combinations of xa, xb, xc with xa + xb + xc = 1 and plot the cloud of points.

```{r}
abc <- read.table("http://www.stat.ucla.edu/~nchristo/datac183c283/statc183c283_abc.txt", 
                header=T)
head(abc)

#expected return
r3 <- abc$a * as.numeric(means[1]) + abc$b * as.numeric(means[4]) + abc$c * as.numeric(means[5])
r3[1:5]

#standard deviation
sd3 <- sqrt(abc$a^2 * covmat[1,1] + abc$b^2 * covmat[4,4] + abc$c^2 * covmat[5,5] + 
              2 * abc$a * abc$c * covmat[1,5] + 2 * abc$a * abc$b * covmat[1,4] + 
              2 * abc$b * abc$c * covmat[4,5])
sd3[1:5]

plot(sd3, r3, xlab = "Risk (standard deviation)", ylab = "Expected Return", 
     main = "Portfolio possibilities curve of Exxon-Mobil, McDonalds, and Boeing")
```

## (f) Assume Rf = 0.001 and that short sales are allowed. Find the composition, expected return and standard deviation of the portfolio of the point of tangency G and draw the tangent to the efficient frontier of question (e).

```{r}
rr <- r[, c(1, 4, 5)]

rf <- 0.001
ri <- colMeans(rr)
R <- ri - rf
R

SIGMA <- cov(rr)
SIGMA

Z <- solve(SIGMA) %*% R
Z

#composition
X <- Z / sum(Z)
X

#expected return
R_G <- t(X) %*% ri
R_G

#standard deviation
SD_G <- sqrt(t(X) %*% SIGMA %*% X)
SD_G

slope <- (R_G - rf) / SD_G
slope

#Find one more point
P <-  rf + slope * 0.3

plot(sd3, r3, xlab = "Risk (standard deviation)", ylab = "Expected Return", 
     main = "Portfolio possibilities curve of Exxon-Mobil, McDonalds, and Boeing")

segments(0, rf, 0.3, P)

#Identify portfolio G:
points(SD_G, R_G, col = "red", pch=19)
text(SD_G, R_G + 0.001, "G")
```

## (g) Find the expected return and standard deviation of the portfolio that consists of 60% and G 40% risk free asset. Show this position on the capital allocation line (CAL).

```{r}
R_Grf <- 0.6 * R_G + 0.4 * rf
R_Grf

SD_Grf <- 0.6 * SD_G
SD_Grf

#Find one more point
P <-  rf + slope * 0.3

plot(sd3, r3, xlab = "Risk (standard deviation)", ylab = "Expected Return", 
     main = "Portfolio possibilities curve of Exxon-Mobil, McDonalds, and Boeing")

segments(0, rf, 0.3, P)

#Identify portfolio G:
points(SD_G, R_G, col = "red", pch=19)
text(SD_G, R_G + 0.001, "G")

#Identify portfolio G and risk free:
points(SD_Grf, R_Grf, col = "green", pch=19)
text(SD_Grf, R_Grf + 0.001, "Grf")
```

## (h) Refer to question (g). Use the expected value (E) you found in (g) to compute x.

```{r}
x <-  (as.numeric(R_Grf - rf) * solve(SIGMA) %*% R) / as.numeric(t(R) %*% solve(SIGMA) %*% R)
x

# x represent the composition of 3 stocks with the expected return of the portfolio of 60% and G 40% risk free asset
```

## (i) (1) Using Rf1 = 0.001 and Rf2 = 0.002 find the composition of two portfolios A and B (tangent to the efficient frontier - you found the one with Rf1 = 0.001 in question (f)).

```{r}
rf2 <- 0.002
ri <- colMeans(rr)
R2 <- ri - rf2
R2

SIGMA <- cov(rr)
SIGMA

Z2 <- solve(SIGMA) %*% R2
Z2

#composition
X2 <- Z2 / sum(Z2)
X2

#expected return
R_G2 <- t(X2) %*% ri
R_G2

#standard deviation
SD_G2 <- sqrt(t(X2) %*% SIGMA %*% X2)
SD_G2

slope2 <- (R_G2 - rf2) / SD_G2
slope2

#Find one more point
P2 <-  rf2 + slope2 * 0.3

plot(sd3, r3, xlab = "Risk (standard deviation)", ylab = "Expected Return", 
     main = "Portfolio possibilities curve of Exxon-Mobil, McDonalds, and Boeing")

segments(0, rf, 0.3, P)
segments(0, rf2, 0.3, P2)

#Identify portfolio G:
points(SD_G, R_G, col = "red", pch=19)
text(SD_G, R_G + 0.001, "G1")

#Identify portfolio G2:
points(SD_G2, R_G2, col = "blue", pch=19)
text(SD_G2, R_G2 + 0.001, "G2")
```

## (i) (2) Compute the covariance between portfolios A and B.

```{r}
cov_ab <- t(X) %*% SIGMA %*% X2
cov_ab
```

## (i) (3) Use your answers to (1) and (2) to trace out the efficient frontier of the stocks Exxon-Mobil, McDonalds, Boeing. Use a different color to show that the frontier is located on top of the cloud of points from question (e).

```{r, warning=FALSE}
set.seed(183)
#many composition
xa <- runif(2000, -1, 3)
xb <- 1 - xa

Rt <- xa * R_G + xb * R_G2
SDt <- sqrt(xa^2 * SD_G^2 + xb^2 * SD_G2^2 + 2 * xa * xb * cov_ab)

plot(sd3, r3, xlab = "Risk (standard deviation)", ylab = "Expected Return", 
     main = "Portfolio possibilities curve of Exxon-Mobil, McDonalds, and Boeing")

#Trace out the efficient frontier: 
points(SDt, Rt, col = "yellow")
```

## (i) (4) Find the composition of the minimum risk portfolio using the three stocks (how much of each stock) and its expected return, and standard deviation.

```{r, warning=FALSE}
#composition
x1_min2 <- (SD_G2^2 - cov_ab) / (SD_G2^2 + SD_G^2 - 2 * cov_ab)
x1_min2
x2_min2 <- 1 - x1_min2
x2_min2

X_1 <- x1_min2 * X[1] + x2_min2 * X2[1]
X_1
X_2 <- x1_min2 * X[2] + x2_min2 * X2[2]
X_2
X_3 <- x1_min2 * X[3] + x2_min2 * X2[3]
X_3


#expected return
r_min2 <- X_1 * as.numeric(ri[1]) + X_2 * as.numeric(ri[2]) + X_3 * as.numeric(ri[3])
r_min2

#standard deviation
sd_min2 <- sqrt(X_1^2 * SIGMA[1,1] + X_2^2 * SIGMA[2,2] + X_3^2 * SIGMA[3,3] + 
              2 * X_1 * X_2 * SIGMA[1,2] + 2 * X_1 * X_3 * SIGMA[1,3] + 
              2 * X_2 * X_3 * SIGMA[2,3])
sd_min2


set.seed(183)
#many composition
xa <- runif(2000, -1, 3)
xb <- 1 - xa

Rt <- xa * R_G + xb * R_G2
SDt <- sqrt(xa^2 * SD_G^2 + xb^2 * SD_G2^2 + 2 * xa * xb * cov_ab)

plot(sd3, r3, xlab = "Risk (standard deviation)", ylab = "Expected Return", 
     main = "Portfolio possibilities curve of Exxon-Mobil, McDonalds, and Boeing")

#Trace out the efficient frontier: 
points(SDt, Rt, col = "yellow")
points(sd_min2, r_min2, col = "red", pch = 19)
```