---
title: "C183 Project 5"
author: "Huimin Zhang"
date: "4/30/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## (a) Assume the single index model holds. Use only the stocks with positive betas in your data. Choose a value of Rf and find the optimal portfolio (point of tangency) using the optimization procedure as discussed in handout #17.

```{r}
#Read csv file:
a <- read.csv("stockData.csv", sep=",", header=TRUE)

#Convert adjusted close prices into returns:
r <- (a[-1,3:ncol(a)]-a[-nrow(a),3:ncol(a)])/a[-nrow(a),3:ncol(a)]

#Compute the betas:
covmat <- var(r)
beta <- covmat[1,-1] / covmat[1,1]

#Keep only the stocks with positive betas:
rrr <- r[, -c(1, which(beta<0)+1)]

#Initialize
beta <- rep(0,ncol(rrr))
alpha <- rep(0,ncol(rrr))
mse <- rep(0,ncol(rrr))
Ribar <- rep(0,ncol(rrr))
Ratio <- rep(0,ncol(rrr))
stock <- rep(0,ncol(rrr))

#Risk free asset:
rf <- 0.001

#This for loop computes the required inputs:
for(i in 1:ncol(rrr)){
	q <- lm(data=rrr, formula=rrr[,i] ~ r[,1])
	beta[i] <- q$coefficients[2] 
	alpha[i] <- q$coefficients[1] 
  mse[i] <- summary(q)$sigma^2
	Ribar[i] <- q$coefficients[1] + q$coefficients[2]*mean(r[,1])
	Ratio[i] <- (Ribar[i]-rf)/beta[i]
  stock[i] <- i
}

#So far we have this table:
xx <- (cbind(stock,alpha, beta, Ribar, mse, Ratio))

#Construct the 30 × 30 variance covariance matrix based on the single index model:
covmat_sim <- covmat[1,1] * (beta %*% t(beta)) + diag(mse)

R <- Ribar - rf

#Compute the vector Z:
Z <- solve(covmat_sim) %*% R

#Compute the vector X:
X <- Z/sum(Z)
head(X)
```

## (b) Rank the stocks based on the excess return to beta ratio and complete the entire table based on handout #37.

```{r}
#Order the table based on the excess return to beta ratio:
A <- xx[order(-xx[,6]),]

col1 <- rep(0,nrow(A))
col2 <- rep(0,nrow(A))
col3 <- rep(0,nrow(A))
col4 <- rep(0,nrow(A))
col5 <- rep(0,nrow(A))

#Create the last 5 columns of the table:
col1 <- (A[,4]-rf) * A[,3] / A[,5]
col3 <- A[,3]^2 / A[,5]
for(i in(1:nrow(A))) {
  col2[i] <- sum(col1[1:i])
  col4[i] <- sum(col3[1:i])
}

#Compute the Ci (col5):
for(i in (1:nrow(A))) {
  col5[i] <- var(r[,1])*col2[i]/(1+var(r[,1])*col4[i])
}

sim_table <- cbind(A, col1, col2, col3, col4, col5)
head(sim_table)
```

## (c) Find the composition of the point of tangency with and without short sales allowed. Place the two portfolios on the plot with the 30 stocks, S&P500, and the efficient frontier that you constructed in the previous projects.

```{r}
#SHORT SALES ALLOWED:
#Compute the Zi:
z_short <- (A[,3]/A[,5]) * (A[,6]-col5[nrow(A)])

#Compute the xi:
x_short <- z_short / sum(z_short)

#The final table when short sales allowed:
Weights_with_short <- cbind(A, col1, col2, col3, col4, col5, z_short, x_short)
Weights_with_short_sort <- Weights_with_short[order(Weights_with_short[,1]),]

#Calculate the expected return and sd of the point of tangency
cov_mat_short <- covmat[1,1] * Weights_with_short_sort[,3] %*% t(Weights_with_short_sort[,3]) + 
  diag(Weights_with_short_sort[,5]) 
sd_short <- (t(Weights_with_short_sort[,13]) %*% cov_mat_short %*% Weights_with_short_sort[,13])^.5
sd_short
R_short <- t(Weights_with_short_sort[,13]) %*% Weights_with_short_sort[,4]
R_short


#SHORT SALES NOT ALLOWED:
#First create a matrix up to the maximum of col5:
table1 <- cbind(A, col1, col2, col3, col4, col5)
table2 <- table1[1:which(col5==max(col5)), ]

#Compute the Zi:
z_no_short <- (table2[,3]/table2[,5]) * (table2[,6]-max(col5))

#Compute the xi:
x_no_short <- z_no_short / sum(z_no_short)

#The final table when short sales are not allowed:
Weights_with_no_short <- cbind(table2, z_no_short, x_no_short)
Weights_with_no_short_sort <- Weights_with_no_short[order(Weights_with_no_short[,1]),]

#Calculate the expected return and sd of the point of tangency
cov_mat_no_short <- covmat[1,1] * Weights_with_no_short_sort[,3] %*% t(Weights_with_no_short_sort[,3]) + 
  diag(Weights_with_no_short_sort[,5]) 
sd_no_short <- (t(Weights_with_no_short_sort[,13]) %*% cov_mat_no_short %*% Weights_with_no_short_sort[,13])^.5
sd_no_short
R_no_short <- t(Weights_with_no_short_sort[,13]) %*% Weights_with_no_short_sort[,4]
R_no_short

## The frontiers using the historical variance covariance matrix (project 2)
#Convert adjusted close prices into returns:
rr <- (a[-1,4:ncol(a)] - a[-nrow(a),4:ncol(a)]) / a[-nrow(a),4:ncol(a)]

#Compute the means:
means.rr <- colMeans(rr)
#Find the covariance matrix:
covmat.rr <- cov(rr)
#Compute the vector of variances:
variances.rr <- diag(covmat.rr)

ones <- matrix(rep(1, 30))

#Compute A:
A <- as.numeric(t(ones) %*% solve(covmat.rr) %*% means.rr)
#Compute B:
B <- as.numeric(t(means.rr) %*% solve(covmat.rr) %*% means.rr)
#Compute C:
C <- as.numeric(t(ones) %*% solve(covmat.rr) %*% ones)
#Compute D:
D <- as.numeric(B*C - A^2)

plot(0, A/C, main = "Portfolio possibilities curve (hyperbola)", type = "n", 
     xlab = "Risk (standard deviation)", ylab = "Expected Return", 
     xlim = c(0, 0.1), ylim = c(-0.04, 0.06))

#Find the asymptotic:
V <- seq(0, 2, 0.001)
A1 <- A/C + V * sqrt(D/C)
A2 <- A/C - V * sqrt(D/C)

#Efficient frontier:
minvar <- 1/C
minE <- A/C
sdeff <- seq((minvar)^0.5, 1, by = 0.0001)
options(warn = -1)
y1 <- minE + sqrt(D*(C*sdeff^2 - 1)) * minvar
y2 <- minE - sqrt(D*(C*sdeff^2 - 1)) * minvar
options(warn = 0)

points(sdeff, y1, type = "l")
points(sdeff, y2, type = "l")

#Compute the means:
means.rr <- colMeans(rr)
#Compute the vector of standard deviations:
stdev.rr <- diag(covmat.rr)^.5

#add the 30 stocks
points(stdev.rr, means.rr)

#add the S&P500
rsp <- (a[-1,3]-a[-nrow(a),3])/a[-nrow(a),3]
meansp <- mean(rsp)
sdsp <- sd(rsp)

points(sdsp, meansp, col = "orange", pch = 19)
text(sdsp, meansp-0.005, "S&P500", col = "orange")

points(sd_short, R_short, col = "blue", pch = 19)
text(sd_short, R_short+0.005, "SIM short sell", col = "blue")

points(sd_no_short, R_no_short, col = "red", pch = 19)
text(sd_no_short, R_no_short+0.005, "SIM no short sell", col = "red")
```

## (d) We want now to draw the efficient frontier when short sale are not allowed.

```{r}
Rfr <- seq(-0.05,.01,0.0005)

#Initialize the two vectors:
rbar_opt <- rep(0,length(Rfr))
risk_opt <- rep(0,length(Rfr))

for(l in 1:length(Rfr)){
#Risk free asset:
rf <- Rfr[l]
#Initialize
beta <- rep(0,ncol(rrr))
alpha <- rep(0,ncol(rrr))
mse <- rep(0,ncol(rrr))
Ribar <- rep(0,ncol(rrr))
Ratio <- rep(0,ncol(rrr))
stocknum <- rep(0,ncol(rrr))

#This for loop computes the required inputs:
for(i in 1:ncol(rrr)){
	q <- lm(data=rrr, formula=rrr[,i] ~ r[,1])
	beta[i] <- q$coefficients[2] 
	alpha[i] <- q$coefficients[1] 
  mse[i] <- summary(q)$sigma^2
	Ribar[i] <- q$coefficients[1] + q$coefficients[2]*mean(r[,1])
	Ratio[i] <- (Ribar[i]-rf)/beta[i]
  stocknum[i] <- i
}

#So far we have this table:
xx <- (data.frame(stocknum,alpha, beta, Ribar, mse, Ratio))

#Order the table based on the excess return to beta ratio:
A <- xx[order(-xx[,6]),]

col1 <- rep(0,nrow(A))
col2 <- rep(0,nrow(A))
col3 <- rep(0,nrow(A))
col4 <- rep(0,nrow(A))
col5 <- rep(0,nrow(A))

#Create the last 5 columns of the table:
col1 <- (A[,4]-rf) * A[,3] / A[,5]
col3 <- A[,3]^2 / A[,5]
for(i in(1:nrow(A))) {
  col2[i] <- sum(col1[1:i])
  col4[i] <- sum(col3[1:i])
}

#Compute the Ci (col5):
for(i in (1:nrow(A))) {
  col5[i] <- var(r[,1])*col2[i] / (1+var(r[,1])*col4[i])
}

#The final table when short sales are not allowed:
B <- cbind(A, col1, col2, col3, col4, col5)
rownames(B) <- NULL
table2 <- B[1:which(col5==max(col5)), ]

#Compute the Zi:
z_no_short <- (table2[,3]/table2[,5])*(table2[,6]-max(col5))

#Compute the xi:
x_no_short <- z_no_short/sum(z_no_short)

#Compute the mean and variance for each portfolio when short sales not allowed:
#First match the columns of the data with the composition of the portfolio:
r1 <- data.frame(rrr[,table2[,1]])

beta1 <- rep(0,ncol(r1))
sigma_e1 <- rep(0,ncol(r1))
alpha1 <- rep(0,ncol(r1))

for(i in 1:ncol(r1)){
	q1<- lm(r1[,i] ~ r[,1])
  beta1[i] <- q1$coefficients[2] 
  sigma_e1[i] <- summary(q1)$sigma^2
  alpha1[i] <- q1$coefficients[1] 
} 

means1 <- colMeans(r1)

#Construct the variance covariance matrix using SIM:
xx <- rep(0,ncol(r1)*(ncol(r1)))                    #Initialize
varcovar <- matrix(xx,nrow=ncol(r1),ncol=ncol(r1))  #the variance covariance matrix 

for (i in 1:ncol(r1)){
	for (j in 1:ncol(r1)){
		varcovar[i,j]=beta1[i]*beta1[j]*var(r[,1])
		if(i==j){varcovar[i,j]=beta1[i]^2*var(r[,1])+ sigma_e1[i]}
	 }
 }
		
rbar_opt[l] <- t(x_no_short) %*% means1
risk_opt[l] <- (t(x_no_short) %*% varcovar %*% x_no_short)^.5
}

plot(risk_opt, rbar_opt, type="l", main="Efficient frontier when short sales not allowed", 
     ylab="Portfolio expected return", xlab="Portfolio standard deviation")
```

## (e) Assume the constant correlation model holds. Rank the stocks based on the excess return to standard deviation ratio and complete the entire table based on handout #44.

```{r}
#Compute the average correlation:
rho <- (sum(cor(r[2:31]))-30)/(30*29)

#Initialize the vectors:
col1 <- rep(0,10)
col2 <- rep(0,10)
col3 <- rep(0,10)

#Compute necessary quantities:
Rbar <- colMeans(r[2:31])
Rbar_f <- Rbar - 0.001
sigma <- (diag(var(r[2:31])))^0.5
Ratio <- Rbar_f/sigma

#Initial table:
xx <- (cbind(Rbar, Rbar_f, sigma, Ratio))

#Order the table based on the excess return to sigma ratio:
aaa <- xx[order(-Ratio),]

#Create the last 3 columns of the table:
for(i in(1:30)) {
  col1[i] <- rho / (1-rho+i*rho)
  col2[i] <- sum(aaa[,4][1:i])
}

#Compute the Ci:
for(i in (1:30)) {
  col3[i] <- col1[i] * col2[i]
}

#Create the entire table until now:
xxx <- cbind(aaa, col1, col2, col3)
head(xxx)
```

## (f) Find the composition of the point of tangency with and without short sales allowed. Place the two portfolios on the plot with the 30 stocks, S&P500, and the efficient frontier that you constructed in the previous projects.

```{r}
#SHORT SALES ALLOWED:
#Compute the Zi:
z <- (1/((1-rho)*xxx[,3]))*(xxx[,4]-xxx[,7][nrow(xxx)])

#Compute the xi:
x <- z/sum(z)

#The final table:
aaaa <- cbind(xxx, z, x)

#SHORT SALES NOT ALLOWED:
#Find composition of optimum portfolio when short sales are not allowed:
aaaaa <- aaaa[1:which(aaaa[,7]==max(aaaa[,7])), ]

#Compute the Zi:
z_no <- (1/((1-rho)*aaaaa[,3]))*(aaaaa[,4]-aaaaa[,7][nrow(aaaaa)])

#Compute the xi:
x_no <- z_no/sum(z_no)

#Final table:
a_no <- cbind(aaaaa, z_no, x_no)

#Initialize the var-covar matrix:
mat <- matrix(rep(0,900), ncol=30, nrow=30)

#Var-covar matrix based on the constant correlation model:
for(i in 1:30){
	for(j in 1:30){
	if(i==j){
		mat[i,j]=aaaa[i,3]^2
		} else {
		  mat[i,j]=rho*aaaa[i,3]*aaaa[j,3]
    }
  }
}

#Calculate the expected return and sd of the point of tangency 
#when short sales allowed
sd_p_opt <- (t(x) %*% mat %*% x)^.5
sd_p_opt
R_p_opt <- t(x) %*% aaaa[,1]
R_p_opt

#Calculate the expected return and sd of the point of tangency 
#when short sales are not allowed
sd_p_opt_no <- (t(x_no) %*% mat[1:which(aaaa[,7]==max(aaaa[,7])),1:which(aaaa[,7]==max(aaaa[,7]))] 
                %*% x_no)^.5
sd_p_opt_no
R_p_opt_no <- t(x_no) %*% aaaaa[,1]
R_p_opt_no

## The frontiers using the historical variance covariance matrix (project 2)
#Convert adjusted close prices into returns:
rr <- (a[-1,4:ncol(a)] - a[-nrow(a),4:ncol(a)]) / a[-nrow(a),4:ncol(a)]

#Compute the means:
means.rr <- colMeans(rr)
#Find the covariance matrix:
covmat.rr <- cov(rr)
#Compute the vector of variances:
variances.rr <- diag(covmat.rr)

ones <- matrix(rep(1, 30))

#Compute A:
A <- as.numeric(t(ones) %*% solve(covmat.rr) %*% means.rr)
#Compute B:
B <- as.numeric(t(means.rr) %*% solve(covmat.rr) %*% means.rr)
#Compute C:
C <- as.numeric(t(ones) %*% solve(covmat.rr) %*% ones)
#Compute D:
D <- as.numeric(B*C - A^2)

plot(0, A/C, main = "Portfolio possibilities curve (hyperbola)", type = "n", 
     xlab = "Risk (standard deviation)", ylab = "Expected Return", 
     xlim = c(0, 0.1), ylim = c(-0.06, 0.08))

#Find the asymptotic:
V <- seq(0, 2, 0.001)
A1 <- A/C + V * sqrt(D/C)
A2 <- A/C - V * sqrt(D/C)

#Efficient frontier:
minvar <- 1/C
minE <- A/C
sdeff <- seq((minvar)^0.5, 1, by = 0.0001)
options(warn = -1)
y1 <- minE + sqrt(D*(C*sdeff^2 - 1)) * minvar
y2 <- minE - sqrt(D*(C*sdeff^2 - 1)) * minvar
options(warn = 0)

points(sdeff, y1, type = "l")
points(sdeff, y2, type = "l")

#Compute the means:
means.rr <- colMeans(rr)
#Compute the vector of standard deviations:
stdev.rr <- diag(covmat.rr)^.5

#add the 30 stocks
points(stdev.rr, means.rr)

#add the S&P500
rsp <- (a[-1,3]-a[-nrow(a),3])/a[-nrow(a),3]
meansp <- mean(rsp)
sdsp <- sd(rsp)

points(sdsp, meansp, col = "orange", pch = 19)
text(sdsp, meansp-0.005, "S&P500", col = "orange")

points(sd_short, R_short, col = "blue", pch = 19)
text(sd_short, R_short+0.005, "SIM short sell", col = "blue")

points(sd_no_short, R_no_short, col = "red", pch = 19)
text(sd_no_short, R_no_short+0.005, "SIM no short sell", col = "red")

points(sd_p_opt, R_p_opt, col="green", pch=19)
text(sd_p_opt, R_p_opt-0.005, "CCM short sell", col = "green")

points(sd_p_opt_no, R_p_opt_no, col="purple", pch=19)
text(sd_p_opt_no, R_p_opt_no-0.005, "CCM no short sell", col = "purple")
```